Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-11-07 paper2_JNLS/paper.tex
NM

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}. More recently, Qi and Majda demonstrated the capability of machine learning strategies to accurately predict these extreme events \cite{MachineLearning2019}. That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}. It should be noted that, in these studies, the extreme events can occur intermittently and on a relatively frequent basis, in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
\caption{
Comparison of surface slope histograms. The experiments and TKdV simulations show very similar slope distributions upstream. Downstream, the theory accounts for the spread of the distribution and the long exponential tails, while differences in the detailed shape of the distributions are also visible. Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements and the associated amplification of noise.}
\label{slopehist}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

>>>>

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}. More recently, Qi and Majda demonstrated the capability of machine learning strategies to accurately predict these extreme events \cite{MachineLearning2019}. That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}. It should be noted that, in these studies, extreme events are represented by strong skewness of the wave-field. These events occur intermittently and on a relatively frequent basis, which is in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
\caption{
Comparison of surface slope histograms. The experiments and TKdV simulations show very similar slope distributions upstream. Downstream, the theory accounts for the spread of the distribution and the long exponential tails, while differences in the detailed shape of the distributions are also visible. Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements.}
\label{slopehist}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
